---
title: "dataType"
author: "Yingrong Chen"
date: "11/21/2021"
output: html_document
---

#Scalar

#Vector
##Key attributes

1. A type, of which there are six major ones:

*Logical
*Character
*Integer
*Double (a way to store non-integer numbers)
NA - meaning the element is empty. It has no value.
NaN - meaning the element has a value, but R doesn’t recognize it as a number. NaN = Not a Number.
Inf - meaning the value is infinity
-Inf - meaning the value is negative infinity
*Complex
*Raw
- Numeric vectors is an umbrella term for integer and double vectors. 
- is.xx() helper functions to check the type of a vector
- typeof(vec) finds the type


2. A length, which is the number of elements it contains

##Missing Values in Vectors
NA refers to the absence of a value for one specific element of a vector
NULL refers to the absence of a vector itself
```{r}
# A logical NA
lgl_vec <- NA

# An integer NA
int_vec <- NA_integer_

# A double NA
dbl_vec <- NA_real_

# A character NA
char_vec <- NA_character_

# A NULL vector
vec_null <- NULL
```

##Subsetting Vectors
```{r}
# Get the first element twice, then all but the last element
bday[c(1, 1, 5, 3)]

# Get everything but the 2nd and 4th elements
bday[c(-2, -4)]

# Find locations of element "Birthday"
which(bday == "Birthday") # Returns a numeric vector of locations of elements meeting the indicated criterion

# Use this to extract all instances of "Birthday" from our vector
bday[which(bday == "Birthday")]

#Return all primes <= 5
five_primes[five_primes <= 5]

#Return all non-missing elements
five_primes[!is.na(five_primes)]
```

#Lists
1.Heterogeneous: an contain more than one type of data
```{r}
# A list
bday_list <- list("Happy Birthday to You!", "How Old Are You?", 34L, TRUE)
str(bday_list)
```
2.Recursive: elements of a list can be vectors themselves or even other lists

##Subsetting Lists
1.single brackets [i] extracts the i’th sub-element of a list and will always return a list of the sub-elements
2.double brackets [[i]] extracts the i’th component of a list and will return whatever that actual things is - a list or a vector of the appropriate type

#Augmented Vectors
vectors with some additional attributes/additional metadata (that is, information) that’s attached to a vector or list that alters its behavior

##attributes():
1. Names
2. Dimensions
turn a vector into a matrix/array.specify a height of vectors via dimension attributes.
3. Classes
specify how objects like vectors and data frames interact with various functions in the S3 object oriented system.

##Types
1. Dates and Date-Times

2. Data Frames and Tibbles
3 new class attributes: tbl, tbl_df, and data.frame
2 new name attributes:names for column (variable) names; row.names to identify observation numbers

3. Factors
Factors represent categorical data, but they’re actually integer vectors with a levels attribute. That attribute assigns a character name to each integer value.