---
title: "Data transformation"
author: "Yingrong Chen"
date: "9/27/2021"
output: html_document
---
dplyr package - part of the tidyverse
## major functions/verbs
filter(): a subset of rows
select(): a subset of variables
arrange(): Reorder the rows of a data frame
mutate(): Create new variables based on existing ones
summarise(): Compute summary statistics
group_by(): Apply any of the above functions to a bunch of sub-groups of the data 

Shared structure:
1. first argument is a data frame or tibble to operate on
2. subsequent arguments tell R what to do to the data frame using variable names (columns) with no quotation marks
3. produce a new data frame

## filter()
choose a subset of rows based on one or more criteria
+ comparison opeators: ==(equal to), != (NOT equal to), >=, <=, >, <
+ Boolean logic: the operators & (for AND, can also use commas , between criteria ), | (“pipe”, for OR), ! (“bang”, for NOT)  
```{r filter}
#Choose flights headed to Albuquerque, NM
# used an unnamed argument since all these verbs have data = as their first argument.
abq_flights <- filter(flights, dest == "ABQ")
#%in%: for a long list of values you want to check a variable against
flights_filtered <- filter(flights, month %in% c(1:6))
#looking for missing NA values, use is.na().
flights_filtered <- filter(flights, is.na(dep_delay))

```

## select()
choose a subset of columns or reorder some columns. select() always gives you the columns in the order you specify
```{r select}
#Select Multiple Adjacent Columns, - for what NOT to Select, names begin with/ conatin "sth."
flights_selected <- select(flights, year:sched_dep_time, -(year:sched_dep_time), starts_with("sched"), contains("dep"), ends_with("dep")) 
#Move 2 variables to be first two columns, then ask for everything() else
flights_selected <- select(flights, carrier, tailnum, everything())
```

## arrange()
re-arranges (or sorts) the rows based on the columns you specify, by default it sorts them in ascending order (lowest-to-highest).
```{r}
flights_arranged <- arrange(flights, sched_dep_time)
flights_arranged <- arrange(flights, desc(sched_dep_time))
#Sort on Multiple Columns
flights_arranged <- arrange(flights, sched_dep_time, dep_time)
```

## mutate()
create new columns 
1. by doing arithmetic operations on existing variables/columns
```{r}
#Create new variable of log(GDP per capita)
gapminder_mutate <- mutate(gapminder, log_gdpPercap = log10(gdpPercap))
# Create a new variable for how far above or below the mean departure delay each flight is
flights_mutate <- mutate(flights_small,
                         relative_delay = dep_delay - mean(dep_delay, na.rm = TRUE))
```

2. lead() and lag()
create a new variable that takes the value of a variable from the next (leading, using lead()) or previous (lagged, using lag()) observation
Caution:
1. Sort data frame properly
2. The first observation in the data will get an NA for a lag(); same for the last observation 
```{r}
# Using lead() and lag() with mutate()
gapminder_mutate <- mutate(gapminder,
#Get population from the next row in the data frame
pop_lead = lead(pop),
#Get population from the previous row in the data frame
pop_lagged = lag(pop))
```

3. Cumulative Functions
(cumulative) sum, mean, product, minimum, or maximum
```{r}
flights_mutate <- mutate(flights_small,
#Cumulative total departure delays across all flights in the data so far
dep_delays_tot = cumsum(dep_delay),
#Cumulative mean departure delay across all flights in the data so far
dep_delays_cummean = cummean(dep_delay))
```

4. If-Else Functions with case_when() and Logical Comparisons
```{r}
flights_mutate <- mutate(flights_small,
#Create new categorical variable for departure depending on values of dep_delay
dep_cat = case_when(dep_delay < -5 ~ "Early", 
dep_delay <= 5 ~ "On Time", # If the first check was FALSE, check if dep_delay meets the second condition
TRUE ~ "Late")) # or any remaining observations just set their evaluation results to TRUE, and assign our remaining value "Late"
```

## group_by()
grouped summarize(), grouped filter(), ungroup()
```{r}
#Most delayed flight each month
flights_small %>% 
  group_by(month) %>% 
  #Find the max non-NA value of `dep_delay`, which will be applied to each month
  filter(dep_delay == max(dep_delay, na.rm = TRUE))
```


## summarize()
summary number (count, mean, sum, maximum, minimum) for the subgroups
median(), quantile(), standard deviation  sd(), interquartile range IQR(), mean absolute deviation  mad(), the number of distinct variable values n_distinct()
```{r}
flights %>% 
  group_by(origin) %>% 
  summarize(mean_dep_delay = mean(dep_delay, na.rm = TRUE),
#asked for the number of observations using n()
  n_flights = n(),
#And the number of flights with non-missing departure delays (only those actually included in our summary)
  n_flights_in_summary = sum(!is.na(dep_delay)))
```

## count()
count the number of observations by groups of one or more variables
```{r}
flights %>% 
  count(origin, month, name = "n_flights")
```

## relational databases
The information you want may be scattered across several data sets. 

### Keys
One or more shared variable(s) that link data sets to one another in pairs  
2 types of keys:
1. Primary keys: variables that uniquely identify a single observation in a table
2. Foreign keys: variables that uniquely identify a single observation in  another table of interest. 

### Relations
A primary key in one data frame and a foreign key in another form a relation.
3 types of relations :
1. 1-to-1: there are single records that match the key in each data frame 
2. 1-to-many: there are single records that match the key in one data frame, but many records that match it in another. For example, there’s only one record for each tailnum in planes, but many such records in flights. In other words, each plane only has one tailnum but takes many flights.
3. Many-to-many: there are multiple records that match the key in each data frame. For example, every airline flies to multiple airports, and every airport has multiple airlines departing it.

## Joining Relational Data: Mutating Joins

#Tutorial 3.3
