---
title: "Data importing, exploring and exporting"
author: "Yingrong Chen"
date: "9/26/2021"
output: html_document
---
##Working Directories
default place where R and RStudio will look for or try to write files to
```{r wd}
#Find out what your working directory is
getwd()
#Set the working directory to be the location of the script or R Markdown
setwd()
# or going to the top menu and clicking  Session > Set Working Directory... > To Source File Location.
```

```{r}
# Specify the actual location of the file. Save the CSV file to your working directory or a sub-directory thereof.
  diabobes <- read_csv("[FILE PATH TO DOWNLOADS FOLDER]/obesity-diabetes.csv")
# go to your working directory - indicated by the . - then the folder datasets, then search there for obesity-diabetes.csv
  diabobes <- read_csv("./datasets/obesity-diabetes.csv")
#move up one level by starting the file path with ..
  diabobes <- read_csv("../obesity-diabetes.csv")

```

## Importing
Several ways you can get data:
1. Data that’s native to R (for example, cars) 
2. Data that come from packages: load the package and then call the data frame
```{r import fromm packages}
library(palmerpenguins)
data(penguins) # function to open datasets part of a package
```
3. “flat” data file from local computer or a shared cloud folder (comma-separated value files .csv or Excel files .xls or .xlsx)

```{r data importing}
readr::read_csv()
readxl::read_excel()
diabobes <- read.csv("./obesity-diabetes.csv")
#grab the data on any sheet with sheet = [SHEET NAME] argument
diabobes_trend <- read.excel("./obesity-diabetes-excel.xlsx", sheet = "diabobes_trends")
# for REALLY big file, to speed up importing
data.table::fread()
diabobes <- fread("./obesity-diabetes.csv")
```

4. Querying SQL or other “relational” databases straight from R or Querying data from an application programming interface (API), commonly in hierarchical JSON or .xml formats

###Removing Files You No Longer Need
rm([name of the object, or objects separated by commas]) 

## Exploring
truncated/stored as a tibble.
```{r data exploration}
#Print the entire data frame
penguins
#open up a new window of table
View(penguins)
#Print the first or last few observations
head(penguins, n = 8)
tail(penguins, n = 4)
#Look at the structure of the data frame
str(penguins)
#Get a summary of each variable of the data frame: 
#For numeric variables (doubles and integers), 6-number summary (minimum, maximum, 1st and 3rd quartiles, median, and mean). For categorical variables (factors), the levels of the factor and the number of observations in each.
summary(penguins)
#skim() the data frame: better, more comprehensive
skimr::skim(penguins)
```

## Exporting

```{r}
# the name of the object, the location to save it, the name of the exported file
write_csv(diabobes, "./obesity-diabetes-exported.csv")
# for REALLY big file, to speed up exporting
data.table::fwrite()
fwrite(diabobes, "./obesity-diabetes-exported.csv")
```



