---
title: "Factors"
author: "Yingrong Chen"
date: "11/20/2021"
output: html_document
---

## Introduction
work with categorical variables, variables with a fixed and known set of possible values

Example: recording month, string is not efficient enough

forcats package in the core tidyverse
```{r}
library(tidyverse)
```

## creating a list of the valid levels:
```{r}
month_levels <- c(
  "Jan", "Feb", "Mar", "Apr", "May", "Jun", 
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)
```

```{r}
x1 <- c("Dec", "Apr", "Jan", "Mar")
y1 <- factor(x1, levels = month_levels)
```

1. sort(y1) based on level
2. any values not in the level will be silently converted to NA

```{r}
# Create a factor with levels taken from the data in alphabetical order
factor(x1)
# access the set of valid levels directly
levels(x1)
```

##changing the levels' orders
### fct_reorder()
reordering the levels for factors whose levels are arbitrarily ordered
x, a numeric vector that you want to use to reorder the levels.
Optionally, fun, a function that’s used if there are multiple values of x for each value of f. The default value is median.

### fct_relevel()
any number of levels that you want to move to the front of the line

### fct_reorder2()
colouring the lines on a plot.
reorders the factor by the y values associated with the largest x values

### fct_infreq()
order levels in increasing frequency

### fct_rev()

## changing the levels' values
### fct_recode()
To combine groups, you can assign multiple old levels to the same new level

### fct_collapse()
use e a vector of old levels to combine to new variable

### fct_lump()
lump together all the small groups
1. default: progressively lump together the smallest groups, ensuring that the aggregate is still the smallest group
2. use the n parameter to specify how many groups we want to keep

## Example
```{r}
relig_summary <- gss_cat %>%
  group_by(relig) %>%
  summarise(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )

rincome_summary <- gss_cat %>%
  group_by(rincome) %>%
  summarise(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )

gss_cat %>%
  mutate(marital = marital %>% fct_infreq() %>% fct_rev()) %>%
  ggplot(aes(marital)) +
    geom_bar()
```

