---
title: "Ggplot"
author: "Yingrong Chen"
date: "9/26/2021"
output: html_document
---
---
title: "ggplot"
author: "Yingrong Chen"
date: "9/26/2021"
output: html_document
---
# Data Visualization 
## ggplot2 
“gg” stands for “Grammar of Graphics,” a unified framework consisting of 7 elements:

```{r ggplot, warning=FALSE}
ggplot() #lays down a default (Cartesian) coordinate system

ggplot(data = <DATA>) + # 1. dataset, + at the end of each line to tell ggplot() more layers or other information is coming
  
  <GEOM_FUNCTION>( # 2. geomm - chart type, E.x. geom_histogram, geom_density, geom_jitter
    
     mapping = aes(<MAPPINGS>), # 3. Aesthetic mappings - X and Y, colors, sizes, fonts etc.
     
     stat = <STAT>, # 4. statistical transformation, or "stat" - often default
     position = <POSITION> # 5. Positional adjustments - bar charts stacked or side-by-side?
  ) +
  <COORDINATE_FUNCTION> + # 6. Coordinate system - Cartesian, polar, or maps for spatial data
  <FACET_FUNCTION> # 7. Facetting - split one plot into many sub-plots
```

### Geom - chart type
#### Visualizing Single Variables
Continuous Variables: 
Histograms (geom_histogram), Density Plots (geom_density), 
Frequency Polygons (geom_freqpoly), 
Boxplots (geom_boxplot)
Raincloud Plots: geom_density() + geom_jitter()

Categorical Variables: Bar Charts

#### Visualizing Two (or Three) Variables
#####Two Continuous Variables
Scatterplots: geom_point, geom_jitter, Line: geom_line()

Linear regression: geom_smooth (method = , linetype =, se = TRUE #CI)
1. method: 
“lm” stands for linear model, specify by formula =
"loess" stands for “Locally Estimated Scatterplot Smoothing”, more flexible curves that are more “responsive” to the data,
"gam" stands for “Generalized Additive Model”, need to specify formula = s(x, bs = "cs")
"glm" stands for “generalized linear model.”

#####One Continuous and One Categorical Variable
Frequency Polygons, Histograms, Density Plots, Boxplots, Raincloud Plots, Ridgeplots (Joyplots): geom_density_ridges(), 

#####Two Categorical Variables
Bar Charts
Heat Maps
```{r}
ggplot(data = count(diamonds, cut, color), mapping = aes(x = color, y = cut)) +
  geom_tile(mapping = aes(fill = n))
```

### Aesthetic Mappings (Aesthetics)
1. The first argument: mapping, which tells the geom layer which aesthetics to associate with which values of the data.
2. Inside mapping = aes()： map to a specific variable in your data，E.g. x to map to flipper_length_mm.
grouping aesthetic - based on one variable:
fill = the color of shapes 
color = the color of points, lines, and outlines (based on one variable
Shape
Size
Alpha (transparency): most useful for overlapping data


3. Outside mapping = aes(): manually-defined aesthetics. 
color = "darkgreen", shape = 15 (15 in 25 built-in shapes for points)

4. Global vs. Local Aesthetics
when we have two geoms (chart type layers), we can make aesthetics global rather than local to avoid typing them twice.
```{r aesthetics}
ggplot(data = penguins) +
  geom_density(mapping = aes(x = flipper_length_mm)) +
  geom_jitter(mapping = aes(x = flipper_length_mm, y = 0), height = 0.001, width = 0)
```

### Statistical Transformations (Stats)
do something to the data before mapping to specific aesthetics
1. stat = "identity", “do nothing” - map the data just as they are (identically), default of geom_point()
2. stat = "count", count the number of observations at each leveland plot that, rather than any actual value from the dataset, works with categorical variables, for bar plot
3. stat = "bin" to “bin” your data in different groups and then plot bin counts, for histograms and frequency polygons
4. stat = "density" to compute the density values, for density plot
5. stat = "smooth" to fit a smoothed curve to your data, geom_smooth()
6. stat = "boxplot" to compute a 5-number summary (min, 1st quartile, median, 3rd quartile, max), boxplots

### position
1. position = "identity" #overlapping
2. position = "stack" #stacking, the default for a bar chart
3. position = "fill" #makes all the bars the same height - 1.0
4. position = "dodge" #side-by-side bars for each color within each cut

### Facetting
break a single plot into sub-plots by groups, usually defined by a categorical variable.
1. facet_wrap(~ class, nrow = 2) # The syntax looks a little funny but it just means facet by class, and arrange the resulting facets in 2 rows
2. facet_grid(drv ~ cyl) #drv = front, rear or 4-wheel drive; cyl = number of cylinders in engine
formula notation: create rows of plots from drv and columns from cyl.
3. scales = "free" / scales = free_x / scales = free_y # The default is "fixed", meaning the scales are fixed (the same) across sub-plots

###Coordinate Systems
Cartesian (X-Y) coordinate, polar coordinate, map for spatial data
coord_flip()

###Customizing Your Plots
1. labs(title = "", subtitle = "", x = "", y = "", caption = "")
2. Legend 
```{r}
#Manually change our legend labels (make sure you have the right order!)
  scale_color_discrete(labels=c("Adelie", "Chinstrap (the best!)", "Gentoo")) + #"_discrete" tells ggplot we're dealing with a categorical variable
  scale_shape_discrete(labels=c("Adelie", "Chinstrap (the best!)", "Gentoo")) +
  
  theme(legend.position = "bottom", # Move legend to bottom
        legend.title = element_blank()) # Delete legend title
```
3. Axis Limits
ylim(min, max), xlim()

4. Reordering Factors (categorical variable)
aes(x = reorder(species, bill_length_mm, FUN = median, na.rm = TRUE)
- reorder the factor species, based on median bill length, remove any NA values of bill length first

5. Themes
pre-made settings on background color, grid lines, and font sizes and types
  theme_bw(), theme_economist(), theme_fivethirtyeight()
  
6. Color Palettes
1. “rainbow palette”: default colors of ggplot2 
2. viridis: scale_color_viridis_c(option = "viridis") for continuous (c) variable and scale_color_viridis_d() for categorical(d) variable
3. RColorBrewer package: sequential data, qualitative data, diverging data
  display.brewer.all()
  scale_color_brewer(palette = "Dark2") for color
  scale_fill_brewer() for fill 
  scale_color_distiller(), for continuous variables 

7. Dealing With Overplotting
1. Jittering (geom_jitter(height=,  width) = ) automatically adding random noise to separate points plotted over one another
2. Alpha: doesn't move every point slightly or obscure their exact value in a bin
3. Hexbin Plots: hexbin package, bin scatterplot values into slightly-larger hexagons

### Storing and Exporting ggplot objects
1. assign the plot to an object, then come back to manipulate it later
2. exporting a plot: ggsave()

ggsave(filename = "C:/Users/elzat/Dropbox/QTM_150/Tutorials/0 - Basics/penguin_plot2.png", plot = penguin_plot2, width, heigh