---
title: "If-Else&For-Loops"
author: "Yingrong Chen"
date: "11/21/2021"
output: html_document
---
Conditional Execution
- code should run (“execute”) different ways depending on whether certain conditions are met

#If-Else Statements
if (condition) {
  # code executed when condition is TRUE
} else {
  # code executed when condition is FALSE
}

Sys.time(): a base R function that extracts the date and time from your system (computer) clock at that instant; it stores it as a date-time object of type POSIXct.

between(): a function from the dplyr package that takes numeric vectors as input and returns TRUE or FALSE depending on whether each element is between the two specified numbers, inclusive.

- A Single If Statement
- A Simple If-Else Statement
- If - Else If - Else Statements

Multiple Conditions
operators || for “OR” and && for “AND”. Don’t use a single | or & as these functions are vectorized, and you don’t want a vector - you want to produce a single TRUE or FALSE.

#For-Loops
- code inside the for-loop will run as long as the conditions outlined are met
```{r}
for(i in (1:15)){   # "FOR THIS": define range of iteration and iteration variable
  
  print(i) # "DO THIS": define what to do at each iteration, commonly involving iteration variable
}
```

```{r}
pacman::p_load(tidyverse, palmerpenguins)
data(penguins)

output <- vector(mode = "double", length = 4)  # Create a output vector to store the result

j <- 1 # Create a new index variable j to help store output

for (i in c(3:6)) { # 1. "FOR THIS" sequence columns 3-6 
    output[[j]] <- median(penguins[[i]], na.rm = TRUE) # 2. "DO THIS" body
    names(output)[j] <- colnames(penguins[i])  # Calculate median and name the resulting value the column name, assign to output vector
    j <- j + 1 # Increment out output vector position by 1
}

output
```

#While-Loops
- code runs until some condition stops being met
```{r}
# Custom function to flip a single coin
flip <- function() sample(c("T", "H"), 1)

flips <- 0 # Set number of flips to start at 0
nheads <- 0 # Set number of heads to start at 0

while (nheads < 13) {
  
  if (flip() == "H") { # Run the flip function once and check the result
    nheads <- nheads + 1 # If its heads, increment that counter
  } 
  
  flips <- flips + 1 # Increment the flip counter no matter what
  
}

#Print the number of flips
flips
```

#Creating Your Own Functions
- a function is a reusable piece of code that takes some input (arguments) and returns some output.
- if you find yourself copy-pasting the same code structure more than twice, maybe it should be a function instead. 

```{r}
# Re-scale a single variable and save it to a new column
# From
penguins$body_mass_rescale <- (penguins$body_mass_g - min(penguins$body_mass_g, na.rm = TRUE)) / 
  (max(penguins$body_mass_g, na.rm = TRUE) - min(penguins$body_mass_g, na.rm = TRUE)) 
# To
vec <- penguins$body_mass_g

penguins$body_mass_rescale <- (vec - min(vec, na.rm = TRUE)) / 
  (max(vec, na.rm = TRUE) - min(vec, na.rm = TRUE)) 

# To
rng <- range(vec, na.rm = TRUE)

penguins$body_mass_rescale <- (vec - rng[1]) / 
  (rng[2] - rng[1]) 

# Let's check if the new column ranges 0 to 1
range(penguins$body_mass_rescale, na.rm = TRUE)
```

```{r}
# Make our rescale function more explicit
rescale01 <- function(vec) {
  
  rng <- range(vec, na.rm = TRUE)    
  rescaled_vec <- (vec - rng[1]) / (rng[2] - rng[1]) # Save to an object you can explicitly return 
  
  return(rescaled_vec) # Our function already did this by default since calculating rng was the last thing it did

}
```

##General Approach to Writing Functions
Write some code
Make sure it works for a specific, simple situation
Realize you want to actually make it a function
Try and abstract and generalize it

#Statistical Simulations

