---
title: "Lubridate"
author: "Yingrong Chen"
date: "10/31/2021"
output: html_document
---

```{r lubridate package}
library(tidyverse)
library(lubridate) # work with dates and times in R, not part of core tidyverse
library(nycflights13)
```

## Three types of date/time data

Date <date>.
Time within a days <time>.
Date-time <dttm> /POSIXct: date + time: it uniquely identifies an instant in time (typically to the nearest second)

## Current date or date-time
```{r}
today()
now()
```

## Create a date/time
1. parse strings/number into date-times 
```{r strings}
ymd(20170131) # in the order of year, month, day as “y”, “m”, “d” 
dmy("31-Jan-2017")
ymd_hms("2017-01-31 20:11:59") # hour, minute, second
mdy_hm("01/31/2017 08:01")
ymd(20170131, tz = "UTC") # force the creation of a date-time by supplying a timezone
```

2. combine individual components of the date-time across multiple columns
`make_date()`
`make_datetime()`
```{r}
flights %>% 
  select(year, month, day, hour, minute) %>% 
  mutate(departure = make_datetime(year, month, day, hour, minute))
```

3. switch between a date-time and a date
`as_datetime()`
`as_date()`

## Date-time components
1. Getting components
`year()`, `month()`, `mday()` (day of the month), `yday()` (day of the year), `wday()` (day of the week), `hour()`, `minute()`, and `second()`
`label = TRUE` to return the abbreviated name of the month or day 
`abbr = FALSE` to return the full name

2. round the date to a nearby unit of time 
`floor_date()`, `round_date()`, and `ceiling_date()`

3. Settinng components
`update()`
```{r}
(datetime <- ymd_hms("2016-07-08 12:34:56"))
year(datetime) <- 2020
hour(datetime) <- hour(datetime) + 1
update(datetime, year = 2020, month = 2, mday = 2, hour = 2)
datetime
```

```{r flights_dt}
make_datetime_100 <- function(year, month, day, time) {
  make_datetime(year, month, day, time %/% 100, time %% 100)}
flights_dt <- flights %>% 
  filter(!is.na(dep_time), !is.na(arr_time)) %>% 
  mutate(
    dep_time = make_datetime_100(year, month, day, dep_time),
    arr_time = make_datetime_100(year, month, day, arr_time),
    sched_dep_time = make_datetime_100(year, month, day, sched_dep_time),
    sched_arr_time = make_datetime_100(year, month, day, sched_arr_time)) %>% 
  select(origin, dest, ends_with("delay"), ends_with("time"))
```

## Time spans
durations: an exact number of seconds
periods: human units like weeks and months
intervals: a starting and ending point

### Duration
- durations always record the time span in seconds
```{r}
h_age <- today() - ymd(19791014)
h_age
as.duration(h_age)
dseconds(15)
dminutes(10)
dhours(c(12, 24))
ddays(0:5)
dweeks(3)
dyears(1)
```
- add and multiply durations
```{r}
2 * dyears(1)
dyears(1) + dweeks(12) + dhours(15)
```

### Periods
```{r}
seconds(15)
minutes(10)
hours(c(12, 24))
days(7)
months(1:6)
weeks(3)
years(1)
days(50) + hours(25) + minutes(2)
10 * (months(6) + days(1))
```
- compare period with duration
```{r}
one_pm <- ymd_hms("2016-03-12 13:00:00", tz = "America/New_York")
one_pm + ddays(1) # day as 24 hrs 86400s
one_pm + days(1) # day as human units, including leap year and daylight savings time
```

### Intervals
a duration with a starting point, so you can determine exactly how long it is
```{r}
next_year <- today() + years(1)
years(1) / days(1)
(today() %--% next_year) / ddays(1) # to find out how many periods fall into an interval
(today() %--% next_year) %/% days(1)
```

```{r}
birth %>% 
  mutate(weekday = wday(date, label = TRUE)) %>%
  group_by(weekday) %>%
  mutate(avg_birth = mean(births, na.rm = TRUE), n = n())
arrange(birth, desc(births))
```


