---
title: '7.1 - Advanced Visualization: Maps'
output: html_document
---

## Maps and map_data packages
The maps and map_data packages contain outline-defining coordinate sets for many continents, countries, states/provinces, and counties.
### packages needed:
pacman::p_load(tidyverse, ggmap, maps, mapdata, RColorBrewer, colorRamps)

## Map Data
1. long is coordinates of longitude
2. lat is coordinates of latitude
3. group defines to which shape each coordinate belongs, “lift its pen”
4. order is the order in which you want R to connect the coordinates to form the correct shape.
draw the entire northern border of a map together, rather than jumping back and forth between the north and south
5. region and subregion provide up to two names for each group (shape) drawn on the map.

```{r}
usa <- map_data("usa")
```

## Plotting Maps
```{r}
#U.S. state map

us_state <- map_data("state")

ggplot(us_state, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group), 
               # By default there's no color and the fill is black, so you won't see the state outlines if you don't change them; it fills all the enclosed pattern
               color = "black", fill = "lavender") + 
  
  # This sets a fixed aspect ratio so the graph doesn't stretch if you zoom in/out or resize
  # Here, a 1-unit change in X is the same length as a 1.3-unit change in Y
  # You may need different values at different latitudes to get "familiar"-looking maps
  coord_fixed(1.3)
```

## Subsetting and “Zooming In”: Only plot certain states
```{r}
#Only plot certain states
us_state %>% 
  filter(region %in% c("georgia", "florida", "alabama", "tennessee", "north carolina", "south carolina")) %>% 
  ggplot(aes(x = long, y = lat)) +
    geom_polygon(aes(group = group), 
                 color = "black", fill = "red") + 
    coord_fixed(1.3)


#Specify the limits inside coord_fixed()
ga_county <- map_data("county") %>% 
  filter(region == "georgia")
ga_county %>% 
  ggplot(aes(x = long, y = lat)) +
    geom_polygon(aes(group = group), 
                 color = "black", fill = "lavender") + 
    coord_fixed(ratio = 1.3, xlim = c(-85, -83.8), ylim = c(33.2, 34.2)) +
    theme_nothing()
```

## Adding Data 

### add dots indicating the locations of Atlanta, Georgia and Chicago
```{r}
# Coordinates for cities
cities <- tibble(long = c(-84.3880, -87.6298),
                 lat =  c(33.7490, 41.8781),
                 city = c("Atlanta", "Chicago"))
ggplot(us_state, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group), 
               color = "black", fill = "lavender") + 
  # Add in city data as points on the map
  # Need to specify we want the dots to use a different data frame, but we still map x to "long" and y to "lat"
  geom_point(data = cities, color = "blue", size = 2) +
  coord_fixed(1.3)
```

### map the aesthetics (most commonly the fill color) of its groups, regions, or subregions
```{r}
#U.S. state map
us_state <- map_data("state")

ggplot(us_state, aes(x = long, y = lat, 
                     # "region" corresponds to state
                     fill = region)) +
  geom_polygon(aes(group = group), 
               color = "black") + 
  coord_fixed(1.3) +
  # Removes the legend for the fill color. To see why we did this, try deleting this line!
  guides(fill = FALSE)
```

### map with another outside variable
#Link diabetes prevalence data to map data
```{r}
#Read in diabetes data
diabobes <- read.csv("./datasets/obesity-diabetes.csv")
diabobes <- diabobes %>% 
  
#Make all state names lower case to match with map data
  mutate(state = str_to_lower(state))
  
us_state_diab <- us_state %>% 
  left_join(diabobes, by = c("region" = "state")) # Region in map data corresponds to state in diabetes data

#Plot diabetes by state
ggplot(us_state_diab, aes(x = long, y = lat, 
                     fill = diab)) +
  geom_polygon(aes(group = group), 
               color = "black") + 
  coord_fixed(1.3)
```