---
title: "7.2 - Advanced Visualization: Interactive Graphs and Plotly"
output: html_document
---

```{r}
pacman::p_load(tidyverse, plotly, gapminder, RColorBrewer)
```

## plotly vs. ggplot
1. the pipe %>%, never the + operator

2. trace where ggplot() would use geom or layer
create new layers/traces on charts using add_XX() rather than geom_XX(). 
For example, to add a scatterplot player you use add_markers() rather than geom_point()

3. specifying aesthetics you no longer need to place them inside aes(), but you do need to precede the variable name with ~

4. uses layout() rather than theme() to customize many elements of a plot such as the legend

5. plot_ly() doesn’t have an explicit group = aesthetic. You have to either use group_by() on a data frame before calling plot_ly() OR use a grouping aesthetic like color = so plot_ly() will implicitly create the groups for you (SEE the section on line plots below)

6. plot_ly() has different defaults for its geom_bar() equivalent

```{r cars}
# Gapminder scatterplot
plot_ly(data = gapminder, # First specify data = <DATA FRAME NAME>
  
  x = ~ gdpPercap, # Specify the x aesthetic maps to ~ <VARIABLE NAME> (in this case, gdpPercap)
              
  y = ~ lifeExp,  # Specify the y aesthetic maps to ~ <VARIABLE NAME> (in this case, lifeExp)
  color = ~ continent, 
  colors = "Dark2") %>% # Use the "Dark2" ColorBrewer palette
  add_lines() # This is what you use to make a line plot
```

```{r}
  colors = "#1f77b4") %>% # 1. This is the hexadecimal code for plot_ly()'s default muted blue
  add_lines(size = I(0.3)) %>% # 3. We have to use I(0.3) to make R/plot_ly() evaluate it as a single number 0.3 rather than mapping to a variable.
  layout(showlegend = FALSE) # 2. Remove the legend
```

## Histograms
  add_histogram()
  
## Bar Charts
add_bars() works like geom_bar() with stat = "identity"
map the height or length of the bar not to a count, but to a value in the data frame by default. Thus it requires you to specify what value that is as a y aesthetic.
## Specifying Traces: add_XX() vs. add_trace()
1. Using add_XX(), where XX is the trace type you want to add

2. Using the more general and flexible add_trace() with type = and possibly mode = arguments to indicate the type of trace you want to add

## Specifying Multiple mode =s
three possible modes: lines, markers, and text. But that means there are actually 7 different types of charts you can produce with any combination of 1, 2, or all 3 of these
```{r}
add_trace(type = "scatter", # In plot_ly() speak, this line plot is actually of type "scatter" but with "lines" instead of "markers" (points) 
            mode = "lines+markers") %>% 
            size = I(0.3) %>% 
  layout(showlegend = FALSE) 
```

## Highlighting and Subsetting Data by Users
1. Highlighting With Legends
e.g. split by continent
2. Highlighting With the Data Itself

We had to pass gapminder through a function called highlight_key(~ <VARIABLE NAME>). This functions sort of like group_by(); it sets the subgroups of data your user can select.
CRITICAL: This must be the last thing you write just before passing data to plot_ly()

## Modifying Tooltip Hovertext
```{r}
  add_markers(color = I("black"),
               text = ~paste0('</br>County: ', county, # Display "County: <COUNTY NAME>" on one line, then 
                              '</br>State: ', state),  # "State: <STATE NAME>" on the next
              hoverinfo = "text") # ONLY display the info we specify in text =
```

## Animation
```{r}
 add_markers(frame = ~ year, # Create one "frame" of the plot for each value of year
              ids = ~ country) %>%   # Causes points to transition from frame to frame within countries 
                                     # In other words, Kuwait 1952 slides to Kuwait 1957, and so on
  layout(showlegend = FALSE) 
```