---
title: "purr"
author: "Yingrong Chen"
date: "11/21/2021"
output: html_document
---
#Ways to accomplish iteration in R:
1. Copying and pasting
2. For-loops
3. The base R family of apply() functions (including lapply(), vapply(), sapply(), etc.). 
4. `map_*()` functions from the `purrr` package, which is part of the tidyverse
5. Other built-in tidyverse functions such as `rowwise()`, `across()`, `c_across()`

#Data type review
A vector: storing many individual elements (a single number or a single character or string) of the same type in a single object

A data frame: storing many vectors of the same length but possibly of different types in a single object

A list: storing many objects of any type (e.g. data frames, plots, vectors) together in a single object

#Map functions
A map function is one that applies the same action/function to every element of an object (e.g. each entry of a list or a vector, or each of the columns of a data frame).
- the second argument, .f, the function to apply, can be a formula, a character vector, or an integer vector.

##Output: different types 
map(.x, .f) is the main mapping function and returns a list

map_df(.x, .f) returns a data frame

map_dbl(.x, .f) returns a numeric (double) vector

map_chr(.x, .f) returns a character vector

map_lgl(.x, .f) returns a logical vector

##Input
vector (of any type): iteration is done over the entries of the vector

list: iteration is performed over the elements of the list

data frame: iteration is performed over the columns of the data frame.

##Working with vector
###Iterate through the vector c(1, 4, 7) by adding 10 to each entry
```{r}
pacman::p_load(knitr, purrr, tidyverse)
#Defining the function
addTen <- function(.x) { 
  return(.x + 10)
}
#Iterating
map(.x = c(1, 4, 7), 
    .f = addTen)
map(c(1, 4, 7), addTen)
#Returning a data frame
map_df(c(1, 4, 7), function(.x) {
  return(data.frame(old_number = .x, # consistent column names
                    new_number = addTen(.x)))
})
```

###Tilde-dot shorthand for functions
```{r}
#Replacing
function(x) {
  x + 10
}
#with
~{.x + 10}
```
“anonymous” function/a temporary function (that you define as the function argument to the map)
~ : started an anonymous function
.x (or simply .): the argument of the anonymous function

###modify() 
just like the map functions, but always returns an object the same type as the input object
modify_if(), only applies the function to elements that satisfy a specific criteria (specified by a “predicate function”, the second argument called .p)
```{r}
modify_if(.x = list(1, 4, 7), 
          .p = function(x) x > 5,
          .f = addTen)
```

##Working with data frame
```{r}
# to download the data directly:
gapminder_orig <- read.csv("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder-FiveYearData.csv")
# define a copy of the original dataset that we will clean and play with 
gapminder <- gapminder_orig

# apply the class() function to each column into a character vector
gapminder %>% map_chr(class)

# apply the n_distinct() function to each column into a numeric vector
gapminder %>% map_dbl(n_distinct)

#anonymous function that contains class() and n_distinct(), and then combines them into a two-column data frame
gapminder %>% map_df(~(data.frame(n_distinct = n_distinct(.x),
                                  class = class(.x))),
                     .id = "variable") #include variable names as a column named "variable"
```

###map2(): multiple input objects
map2(.x = object1, # the first object to iterate over
     .y = object2, # the second object to iterate over
     .f = plotFunction(.x, .y))
nth iteration will correspond to the nth element in the object1 and the nth element in the object2

e.g. if you have a continent vector .x = c("Americas", "Asia") and a year vector .y = c(1952, 2007), the iteration will be first the Americas for 1952, and then Asia for 2007

###pmap() 
iterate over an arbitrary number of objects (i.e. more than two).

###invoke_map()
The first argument is a list of functions or character vector of function names. The second argument is a list of lists giving the arguments that vary for each function. The subsequent arguments are passed on to every function.

```{r}
f <- c("runif", "rnorm", "rpois")
param <- list(
  list(min = -1, max = 1), 
  list(sd = 5), 
  list(lambda = 10)
)

invoke_map(f, param, n = 5) %>% str()
```


##Tibbles
tibble columns can be lists (as opposed to vectors
a tibble can be “nested” where the tibble is essentially split into separate data frames based on a grouping variable, and these separate data frames are stored as entries of a list
```{r}
gapminder_nested <- gapminder %>% 
  group_by(continent) %>% 
  nest()
gapminder_nested
gapminder_nested$data[[1]] #the first entry in the data column corresponds to the entire gapminder dataset for Asia

```

To apply vectorized functions like `mutate` to a list-column, you need to wrap the function you want to apply in a map function.
```{r}
#calculate the average life expectancy within each continent and add it as a new column
gapminder_nested %>% 
  mutate(avg_lifeExp = map_dbl(data, ~{mean(.x$lifeExp)}))

#fit a model separately for each continent
gapminder_nested <- gapminder_nested %>% 
  mutate(lm_obj = map(data, ~lm(lifeExp ~ pop + gdpPercap + year, data = .x)))
gapminder_nested
```

###keep()
keeps elements of a list that satisfy a given condition, much like select_if() selects columns of a data frame 
discard() does the opposite of keep(): discards any elements that satisfy your logical condition.
```{r}
#keeps the gapminder continent data frames that have an average (among the sample of 5 rows) life expectancy of at least 70
gapminder_list %>%
  keep(~{mean(.x$lifeExp) > 70})
```
###reduce()
combines (reduces) all of the elements of a list into a single object by iteratively applying a binary function (a function that takes two inputs)
accumulate() returns the intermediate values.

```{r}
reduce(c(1, 2, 3), sum)
accumulate(c(1, 2, 3), sum)
```

###every(), some(), has_element()
```{r}
#whether every continent has average life expectancy greater than 70
gapminder_list %>% every(~{mean(.x$life) > 70})
#whether some continents have average life expectancy greater than 70
gapminder_list %>% some(~{mean(.x$life) > 70})

list(1, c(2, 5, 1), "a") %>% has_element("a")
```

#Reading in many files at once 
```{r}
library(palmerpenguins)
#splitting the penguins dataset into three csv files, one for each species, named as [NAME OF SPECIES].csv, and save them in the data/ folder.
penguins %>%
  group_by(species) %>%
  group_walk(~ write_csv(.x, paste0("data/", .y$species, ".csv")))

#purrr solution
map_dfr(files, read_csv, .id = "species")

#vroom()
library(vroom)

vroom(files, id = "species")
```

#Dealing with failure
safely(): it takes a function (a verb) and returns a modified version:

result: the original result. If there was an error, this will be NULL.

error: an error object. If the operation was successful, this will be NULL.

possibly(), simpler than safely(), with a default value to return when there is an error
quietly(), instead of capturing errors, it captures printed output, messages, and warnings
