---
title: "Shiny"
author: "Yingrong Chen"
date: "9/29/2021"
output: html_document
---

Every Shiny App is maintained by a computer running R. To share the shiny, you need to set up a web server to run the App 

# Build an APP
1. user interface (UI)
2. server instructions

## App template 
```{r template}
library(shiny)
ui <- fluidPage()
server <- function(input, output) {}
shinyApp(ui = ui, server = server)
```

## inputs and output
Add elements to your app as arguments to fluidPage()
```{r}
 ui <- fluidPage(
  # *Input() functions,
  # *Output() functions
)
```

### input 
1. Buttons, Checkbox, radio buttons, select box, sliders
2. Data input, Data range, File input, Numeric input, Password input, textinput

syntax: function(inputId = "num", label = "Choose a number", ...)
inputId: input name (for internal use), 
label to display, "" for empty character string
input specific argument

- The input value changes whenever a user changes the input.

### output 
dataTableOutput()
imageOutput()
tableOutput()
htmlOutput()
plotOutput()
textOutput()
verbatimTextOutput(): text
uiOutput(): a Shiny UI element

syntax: function(outputId = "hist")
outputId: name to give to the output object

```{r}
sliderInput(inputId = "num",
  label = "Choose a number",
  value = 25, min = 1, max = 100)
plotOutput("hist")
```
 
## server
Tell the server how to assemble inputs into outputs

1. Save objects to display to output$
2. Build objects to display with render*() function
render*() function analogous to the output function, builds reactive output to display in UI
syntax: function({code block that builds the object}) # {} put as many codes as you want to the same block
3. Access input values with input$

- Create reactivity by using Inputs to build rendered Outputs
```{r}
output$hist <- renderPlot({ 
  title <- "100 random normal values"
  hist(rnorm(input$num), main = title) 
})
```

## Share the App

### save the App
One directory with every file the app needs:
• app.R (your script which ends with a call to shinyApp()) 
• datasets, images, css, helper scripts, etc.
2 file apps: (ui.R) + (server.R)

Use the exact name: (app.R) or (ui.R) + (server.R)! 

### Launch the App
Run App: Run in external, viewer pane and pop up winder

### package the App
https://www.shinyapps.io/: free server maintained by RStudio
scalable: assign extra R processers for several users to avoid slow-down



 

```{r sample}
library(shiny)
# user interface (UI)
ui <- fluidPage( 
  headerPanel('Iris k-means clustering'),
  sidebarPanel(
    selectInput('xcol', 'X Variable', names(iris)),
    selectInput('ycol', 'Y Variable', names(iris),
      selected = names(iris)[[2]]),
    numericInput('clusters', 'Cluster count', 3,
      min = 1, max = 9)
  ),
  mainPanel(
    plotOutput('plot1')
  )
)
# server instructions
server <- function(input, output) {

  selectedData <- reactive({
    iris[, c(input$xcol, input$ycol)]
  })

  clusters <- reactive({
    kmeans(selectedData(), input$clusters)
  })

  output$plot1 <- renderPlot({
    par(mar = c(5.1, 4.1, 0, 1))
    plot(selectedData(),
         col = clusters()$cluster,
         pch = 20, cex = 3)
    points(clusters()$centers, pch = 4, cex = 4, lwd = 4)
  })

}
# Combine ui and server
shinyApp(ui = ui, server = server)
```



