---
title: "Strings"
author: "Yingrong Chen"
date: "10/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Basics
- create strings with either single quotes or double quotes
- stored multiple strings in a character vector c("one", "two", "three")
- string vs. a character vector (vector of strings)
- pairs: one stringr function works with a single match, and the other(_all) works with all matches. 
- stringr, package in tidyverse, built on top of stringi

literal single or double quote \“
 "\n", newline,  "\t", tab

## str_c(): Combining strings in a character vector to a string
```{r}
str_length() # the number of characters in a string
# str_c(): combine strings
str_c("x", "y", "z")
str_c("x", "y", sep = ", ") # sep to control how they’re separated
str_c(c("x", "y", "z"), collapse = ", ") # collapse to collapse a vector of strings into string
str_c("|-", str_replace_na(x), "-|") # str_replace_na(): print missing values as "NA"
str_c("prefix-", c("a", "b", "c"), "-suffix") # vectorised, automatically recycles shorter vectors to the same length as the longest
# Objects of length 0 are silently dropped
```

## str_sub(): Subsetting strings
```{r}
x <- c("Apple", "Banana", "Pear")
str_sub(x, 1, 3)
str_sub(x, -3, -1) # negative numbers count backwards from end
str_sub(x, 1, 1) <- str_to_lower(str_sub(x, 1, 1)) # use the assignment form of str_sub() to modify strings
```

## Make edits
Locales: different rules for different languages, ISO 639 language code
```{r}
str_to_upper(c("i", "ı"))
str_to_upper()
str_to_title()
str_sort()
str_order()
```

## Find patterns
```{r}
str_detect() # returns a logical vector the same length as the input
str_locate() # the starting and ending positions of each match
str_count(): # how many matches there are in a string
```

## 
```{r}
str_match()
str_subset()
str_extract()
```

```{r Example}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
has_colour <- str_subset(sentences, colour_match)
matches <- str_extract(has_colour, colour_match)

noun <- "(a|the) ([^ ]+)"

has_noun <- sentences %>%
  str_subset(noun)
has_noun %>% 
  str_extract(noun)
```

## Replace/Remove patterns
```{r}
str_remove()
str_replace()
```

## Splitting
```{r}
sentences %>%
  head(5) %>% 
  str_split(" ")
```

## Patterns regex()
regex metacharacters: $ . | ? * + ( ) [ {
str_view(fruit, "nana") is shorthand for str_view(fruit, regex("nana"))

\d: any digit.
\s: any whitespace (e.g. space, tab, newline).
[abc]: matches a, b, or c.
[^abc]: matches anything except a, b, or c.
[.] 
```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")
str_view(x, ".a.") # . matches any character (except a newline)
str_view(x, "^a") # ^ matches the start of the string
str_view(x, "a$") # $ matches the end of the string
```

Details of the match
multiline = TRUE allows ^ and $ to match the start and end of each line 
ignore_case = TRUE
comments = TRUE
dotall = TRUE allows . to match everything, including \n.

## Repetition
how many times a pattern matches
?: 0 or 1
+: 1 or more
*: 0 or more

specify the number of matches precisely:
{n}: exactly n
{n,}: n or more
{,m}: at most m
{n,m}: between n and m

- Grouping and backreferences
```{r}
str_view(fruit, "(..)\\1", match = TRUE) #  all fruits that have a repeated pair of letters.
```

## other matchinng functions 
fixed() matches the specified sequence of bytes
coll() compares strings using standard collation rules
boundary() matches boundaries

## other uses of regular expressions
apropos() searches all objects available from the global environment, useful when can’t remember the name of the function.
dir() lists all the files in a directory in certain pattern
e.e dir(pattern = "\\.Rmd$")

