---
title: "Tidy data"
output: html_document
---

```{r Packages}
# Load (and, if necessary, first install) packages
pacman::p_load(tidyverse)

# Abortion data from CDC and Alan Guttmacher Institute (AGI) extracted from 
# http://www.johnstonsarchive.net/policy/abortion/index.html#UC
abortion <- read.csv("./datasets/ab.csv")

# Econ data from the World Bank
real_econ <- read.csv("./datasets/real_econ.csv", fileEncoding = 'UTF-8-BOM') #Required to avoid odd character on first column name
tidy_econ <- read.csv("./datasets/tidy_econ.csv")
```

## Data Cleaning
tidy data: fits with the  tidyverse ecosystem - and many other data analytic ecosystems

The 3 Tidy Criteria:

Each variable must have its own column
Each observation must have its own row
Each value must have its own cell
-depends on the definition of "variable" and "observation"

Other data structure:
more efficient ways to store high volumes of data: efficient hierarchical JSON or XML structures. 
“Long” vs. “Wide” Data
wide format: adds columns to save on rows, a wider rectangular data frame 
long (or narrow) format: adds rows to save on columns, a longer/taller rectangular data frame, tidy data
In the long data we had to repeat the information on certain variables more times than in the wide data.

# Functions 
## separate()
For column with multiple pieces of data.

basic structure: 
<old data frame> %>%
separate(<column you wish to separate>, 
         into = c(<names of the new columns created, separated by comma>), 
         sep= <condition or spot at which you separate old column values>)
         
The sep = argument can be one of two basic types: 
1. a character or string of characters at which you want the separation to occur (it will occur each time that character or string or characters is encountered)
2. a number of characters after which you want to create a split (for example, if you enter 5 it will split the column’s values after the first 5 characters). 

```{r}
table3 %>%
separate(rate, # Separate rate
         into = c("cases", "population"), # Split it into two columns called cases and population 
         sep = "/", # Split the values into pieces before the / and after it
         convert = TRUE) # See if you can figure out what this does. Try this code and without it. Look carefully at the results.
```

## unite()
basic structure: 
<old data frame> %>%
separate(“<name of new column after unification>”, 
         (<names of columns you want to unify, separated by commas), 
         sep = <characters, if any, to insert between values from unified columns>)
         
```{r}
table1 %>%
unite("rate", # Create new rate column
         cases, population, # Unite these two into `rate` 
         sep = "/") # Split the values using a "/" in the new column
```

# Switch between long and wide format
## pivot_longer()
when several observations are combined in the same row. 
reshape wide data frame into long (tidy) format by ensuring there is only one (1) observation - defined as data from a single country in a singe year - in each row. previously known as gather()

<wide data frame> %>% 
  pivot_longer(cols = <names of columns across which single observation is split, chosen using "select" syntax>,
               names_to = <"implicit variable represented in the names of cols">,
               values_to = <"the variable represented by the values stored in cols">)
               
```{r warning=TRUE}
# Change `table4a` from wide to long format
table4a %>% 
  pivot_longer(cols = `1999`:`2001`, # Note the select() style syntax
               names_to = "year",
               values_to = "cases")
```

select() syntax

## pivot_wider()
reshape from long-to-wide, when a single observation is split across multiple rows, previously known as spread()
<long data frame> %>% 
  pivot_wider(names_from = <"column to take new column names from">,
              values_from = <"column that holds the values for the new columns">)
```{r}
# Tidy table2 by widening it
table2 %>% 
  pivot_wider(names_from = "type", # Will create two new columns, one for each value of "type"
              values_from = "count") # Will place values from count into those two new columns
```

